---
output:
  word_document: default
  html_document: default
---
# Module 6 - Assignment 1
## Kiser, Stephen 
### Lubridate
```{r Loading in Tidyverse and Lubridate}
library(tidyverse)

library(lubridate)

Appointments <- read_csv("Appointments.csv")

```

### Working with Dates
```{r working_with_dates}
Appointments <- mutate(Appointments, ReservationYear = 2019)

Appointments <- Appointments %>% 
  mutate(
    ReservationDate = make_date(month=ReservationMonth, day=ReservationDay, year=ReservationYear),
    CreationDate = make_date(month = CreationMonth, day = CreationDay, year= CreationYear),
    ReservationSpan = (ReservationDate - CreationDate)
)

```

### Exploratory Analysis
```{r Exploratory Analysis}
summary(Appointments)
Appointments$ReservationSpan <- as.numeric(as.difftime(Appointments$ReservationSpan, units = "days"))
summary(Appointments)

```

a. What is the average time between appointment creation and the actual reservation date?

The average time between appointment creation and actual reservation date is 38.53 days.

b. What was the max time?

The max time between appointment creation and actual reservation date is 207 days.

c. What was the min time? 

The minimum time between appointment creation and actual reservation date is 1 day.

```{r correlation}
cor(Appointments$ReservationSpan,Appointments$Show,method = "pearson")
```
a. Are ReservationSpan and Show highly correlated?

ReservationSpan and Show are not highly correlated.

```{r bar chart}
ggplot(Appointments, aes(x=Show,))+
  geom_bar()
```

a. Based on 0 being no show and 1 being show, do most people make their appointments?

Yes, most people make their appointments.